Effects of Reynolds Number on the Performance and Near-Wake of a Vertical-Axis Cross-Flow Turbine

### 2015-04-17 paper.tex
Write section on static foil coefficients

\subsubsection{Understanding via Static Foil Coefficients}

To help understand---and possibly predict---the $Re$-sensitivity on turbine
performance, a series of foil coefficient datasets were computed with the XFOIL
viscous panel method code.

The turbine torque coefficient $C_T$ can be related to the blade section
chordwise force coefficient $C_c$ by
\begin{equation}
C_T = \frac{C_c c}{2R} \frac{|U_\mathrm{rel}|^2}{U_\infty^2},
\end{equation}
where the blade section chordwise force (for zero preset pitch)
\begin{equation}
C_c = C_l \sin \alpha - C_d \cos \alpha.
\end{equation}

@@

@@

@@

Note that the common static foil performance metrics fail to show Reynolds
number dependence as close as the max geometric torque coefficient. This means
that the turbine kinematics play an important role, even if the exact dynamics
of the flow are not known.

>>>>

\subsubsection{Relation to Static Foil Characteristics}

To help understand---and possibly predict---the $Re$-sensitivity on turbine
performance, a series of static foil coefficient datasets were computed with the
XFOIL viscous panel method code \cite{Drela1989}, implemented as part of the
open-source turbine design software QBlade. Simulations were run with the
default settings and zero Mach number, for the approximate average blade chord
Reynolds numbers encountered in the experiments.

To investigate the effects of the blades' ``virtual camber'' due to their
circular path \cite{Migliore1980}, the XFOIL calculations were performed for
20\% thick foils with 0\% (NACA 0020), 2\% (NACA 2520), and 4\% (NACA 4520)
camber about their half-chord location. A 4\% camber approximates the maximum
distance between the blade chord line and path for the UNH-RVAT and the 2\%
camber takes into account the reduction in virtual flow curvature from the
non-curved inflow velocity by dividing by the tip speed ratio $\lambda=1.9$.

Results from the XFOIL calculations are shown in Figure~\ref{fig:foil-Re-dep},
where values of maximum lift coefficient, minimum drag coefficient, and maximum
lift-to-drag ratio are normalized (for comparison between foils) and plotted
versus $Re_c$. In general, larger camber is associated with decreased foil
performance at lower Reynolds number.

@@

In conjunction with the cross-flow turbine blade kinematics, the foil
coefficients were used to approximate turbine performance by calculating the
peak torque coefficient on the upstream half of the blade path. The turbine
torque coefficient $C_T$ can be related to the blade section chordwise force
coefficient $C_c$ by
\begin{equation}
C_T = \frac{C_c c}{2R} \frac{|U_\mathrm{rel}|^2}{U_\infty^2},
\end{equation}
where the blade section chordwise force (for zero preset pitch)
\begin{equation}
C_c = C_l \sin \alpha - C_d \cos \alpha.
\end{equation}

The relative blade velocity was calculated by adding the free stream to the
opposite of the blade tangential velocity. Note that this neglects any
induction, or slowing of the free stream by the turbine forces, which would be
present in a momentum/streamtube model. Since we were not looking to predict
absolute performance but rather relative changes with $Re$, this method was
deemed acceptable as it is extremely simple and fast to compute.

Values for the blade angle of attack, relative velocity, and torque coefficient
are plotted in Figure~\ref{fig:blade-kinematics}. The effects of static stall
are clearly present in the torque coefficient plot, and by the time the angle of
attack has decreased below that of stall, the relative velocity is so low that
there is not much contribution to the torque coefficient.

@@

Results for the peak torque coefficient for each foil are plotted versus $Re_c$
in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the convergence of
$C_{T_\mathrm{max}}$ with increasing Reynolds number is more dramatic than any of the
common foil characteristics plotted in Figure~\ref{fig:foil-Re-dep}, meaning
that the cross-flow turbine's unique kinematics must be taken into account when
attempting to predict the effect of transitional Reynolds numbers on turbine
performance.

From the peak torque coefficient metric, the Reynolds number independence is
achieved at lower values, and more dramatically. We see that the trend of the
NACA 0020 curve matches almost perfectly at low $Re$, but the trend to continue
increasing linearly is not matched by the experimental data, which looks to
behave more like the cambered foil results. Though this method is very
approximate, it nevertheless appears to be a reasonable way to predict the
transitional Reynolds number regime for cross-flow turbine performance using
only 2-D static airfoil characteristics

@@
